### 2014-11-17 sharp-sigma-plus.tex
f-1 avance

\begin{theo}
$\#\Sigma_1$ is closed under substraction of 1 element.
\end{theo}
\begin{proof}
The counting set has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
\begin{enumerate}
\item The counting set is:
$$
F(\A) = \vert \{ <\!\!\vec{z}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x},\vec{z}) \} \vert
$$
our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. Let $F'$ be:
$$
F'(\A) = \vert \{ <\!\!\vec{z}\!\!> :\A \models \exists \vec{x}\left( \varphi(\vec{x},\vec{z}) \wedge \exists \vec{z'} (\varphi(\vec{x},\vec{z'}) \wedge \varphi_{\lneq}(\vec{z'},\vec{z} ) ) \right) \} \vert
$$
where $\varphi_{\lneq}(\vec{z'},\vec{z})$ is a formula that indicates that $\vec{z'}$ is lexicographically smaller than $\vec{z}$. Thus, $F'(\A)$ will count exactly one element less than $F(\A)$ if $F(\A)>0$.
\item The counting set is:
$$
F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x},\vec{P}) \} \vert
$$
similarly, we would like to eliminate the lexicographically smallest predicate $P$ that holds the formula true with the exact number of tuples. First we treat quantified sub-formulas as literals, and take a DNF formula that's equivalent to $\varphi$. Then
$$
F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
\exists \vec{x} \varphi_1(\vec{x},\vec{P}) \vee
\exists \vec{x} \varphi_2(\vec{x},\vec{P}) \vee
... \vee
\exists \vec{x} \varphi_k(\vec{x},\vec{P})
 \} \vert
$$
Note that
$$
\varphi_i(\vec{x},\vec{P}) = \vec{P_1}(\vec{x}) \wedge \neg \vec{P_0}(\vec{x}) \wedge \varphi'_i(\vec{x})
$$
where $\vec{P_0},\vec{P_1}$ are subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified. We can assume every predicate in $\vec{P},$ uses all variables in $\vec{x}$ and they are not permutated(?). Let $\psi_i(x)$ be:
$$
\psi_i(\vec{x}) = \exists x' (x' \neq x \wedge P_1(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x))
$$
then 
$$
\exists \vec{x} \left( \varphi_i(\vec{x},\vec{P}) \wedge \psi_i(\vec{x}) \right)
$$
will have exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$.
\end{enumerate}
\end{proof}

>>>>

\begin{theo}
$\#\Sigma_1$ is closed under substraction of 1 element.
\end{theo}
\begin{proof}
The counting set has three possible ways of counting. Counting only variables, only predicates, and predicates with variables. This separates the proof in three cases:
\begin{enumerate}
\item The counting set is:
$$
F(\A) = \vert \{ <\!\!\vec{z}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x},\vec{z}) \} \vert
$$
our goal here is to eliminate the lexicographically smallest sequence of variables, which we can do easily. Let $F'$ be:
$$
F'(\A) = \vert \{ <\!\!\vec{z}\!\!> :\A \models \exists \vec{x}\left( \varphi(\vec{x},\vec{z}) \wedge \exists \vec{z'} (\varphi(\vec{x},\vec{z'}) \wedge \varphi_{\lneq}(\vec{z'},\vec{z} ) ) \right) \} \vert
$$
where $\varphi_{\lneq}(\vec{z'},\vec{z})$ is a formula that indicates that $\vec{z'}$ is lexicographically smaller than $\vec{z}$. Thus, $F'(\A)$ will count exactly one element less than $F(\A)$ if $F(\A)>0$.
\item The counting set is:
$$
F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models \exists \vec{x} \varphi(\vec{x},\vec{P}) \} \vert
$$
similarly, we would like to eliminate the lexicographically smallest predicate $P$ that holds the formula true with the exact number of tuples. First we treat sub-formulas without $\vec{P}$ as literals, and take a DNF formula that's equivalent to $\varphi$. Then
$$
F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
\exists \vec{x} \varphi_1(\vec{x},\vec{P}) \vee
\exists \vec{x} \varphi_2(\vec{x},\vec{P}) \vee
... \vee
\exists \vec{x} \varphi_k(\vec{x},\vec{P})
 \} \vert
$$
Note that
$$
\varphi_i(\vec{x},\vec{P}) = \vec{P_1}(\vec{x}) \wedge \neg \vec{P_0}(\vec{x}) \wedge \varphi'_i(\vec{x})
$$
where $\vec{P_0},\vec{P_1}$ are disjunct subsequences of $\vec{P}$ and $\varphi'_i$ is arbitrarily quantified. We can assume every predicate in $\vec{P}$ uses all variables in $\vec{x}$ and they are not permutated \footnote[1]{revisar esto}. Hence, if there is at least one assignation to $\vec{x}$ that satisfies $\varphi'_i(\vec{x})$, there is a predicate that has exactly one tuple that is lexicographically minimal. Thus, if
$$
\psi_i(\vec{x}) = \exists x' (x' \neq x \wedge P_1(x')) \vee \exists x' (\varphi'_i(x') \wedge \varphi_{\lneq}(x',x))
$$
then 
$$
\exists \vec{x} \left( \varphi_i(\vec{x},\vec{P}) \wedge \psi_i(\vec{x}) \right)
$$
will be satisfied by exactly one predicate less than $\exists \vec{x} \varphi_1(\vec{x},\vec{P})$. If $\varphi'_i(\vec{x})$ is unsatisfiable, we eliminate one predicate on the next $\varphi_i$. If this sub-formula is also unsatisfiable, we move to the next, and so on. Therefore:
\begin{multline*}
F(\A) = \vert \{ <\!\!\vec{P}\!\!> :\A \models 
\exists \vec{x} \left( \varphi_1(\vec{x},\vec{P}) \wedge \psi_1(\vec{x}) \right) \vee
\exists \vec{x} \left( \varphi_2(\vec{x},\vec{P}) \wedge \left( (\forall \vec{x'} \neg \varphi'_1(\vec{x'})) \rightarrow \psi_2(\vec{x}) \right)\right) \vee \\
... \vee
\exists \vec{x} \left( \varphi_k(\vec{x},\vec{P}) \wedge \left( (\forall \vec{x'} (\neg \varphi'_1(\vec{x'}) \wedge \neg \varphi'_2(\vec{x'}) \wedge ... \wedge \neg \varphi'_{k-1}(\vec{x'})) ) \rightarrow \psi_k(\vec{x}) \right)\right)
 \} \vert
\end{multline}
\end{enumerate}
\end{proof}
